% discussion.tex
% Author: Tony Kabilan Okeke
% Date: 2023-03-23

\section{Discussion}

Our results provide insights into the use of autoencoders for unsupervised feature
learning from gene expression data. The performance of the autoencoder was evaluated
based on its ability to reconstruct the input data, and we believe that this is model
captures the underlying structure of the data well. However, we acknowledge some
limitations of our study. One major limitation was the memory constraints we faced
during training, which limited our ability to use all 48978 genes. In addition, we were
only able to evaluate a limited number of hyperparameters due to computational
limitations. Future studies could explore more advanced architectures such as sparse and
variational autoencoders to improve the performance of the model. Despite these
limitations, our study provides a foundation for future work in this area.
We plan to explore the use of the latent space of the autoencoder to perform feature
extraction for a classifier to predict the enrichment of Gene Ontology terms.

\vspace{0.2cm}

In summary, our work highlights the potential of autoencoders for unsupervised feature
learning from gene expression data, and provides a starting point for future studies to
explore more advanced architectures and evaluate the performance of the model on
predicting biological annotations. However, we acknowledge the limitations of our study
and the need for further research to address these limitations and advance the field.
